\forestclaw: Hybrid forest-of-octrees AMR for hyperbolic conservation laws

### 2013-07-30 docs/parco13/paper.tex
Clarify refinement procedure

\newcommand{\comment}[1]{\textcolor{green}{[DAC: #1]}\xspace}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}\xspace}
\newcommand{\sR}{\mathbb{R}}
\newcommand{\dt}{\mathrm{d}t}
\newcommand{\clawpack}{{\sc Clawpack}\xspace}
\newcommand{\forestclaw}{Forestclaw\xspace}
\newcommand{\pforest}{\texttt{p4est}\xspace}
\newcommand{\manyclaw}{Manyclaw\xspace}

@@

A main challenge is presented by the fact that the patch neighborhood
is only known to \pforest.  This patch connectivity information needs
to be propagated to the numerical code in \forestclaw that implements
the interaction with neighbor patches via the use of a layer of ghost
cells surrounding each patch.  \alert{To this end}, we define an interface
that allows read-only access to the sequence of blocks,
the list of patches for each, and a lookup of neighbor patches (and
their relative orientation which is nontrivial between blocks).
Suitably informed by \pforest, \forestclaw stores only the patches local to
each MPI rank.  \alert{The exchange of ghost} data and parallel repartitioning is
provided transparently by \pforest.  Mesh modification directives, such as
adaptive refinement and coarsening, are called from \forestclaw and relayed to
\pforest.

\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
%
Quote: How is AMR working in detail. For example: Figure 1 shows a
2:1 refinement, thus refinement of a single patch could affect a whole region
of the computational domain. + The underlying equations are known to
Forestclaw: How is is the information shared with p4est (equation dependent
refinement criteria, refinement and coarsening strategies (interpolation,
integration).
}

@@

\forestclaw uses iterators over all rank-local leaves, optionally restricted to
a given level.  Random access is possible and used when executing $O(1)$ time
neighbor lookups.  Looping over the patches in the order prescribed by the
forest
\alert{Review 2 - Explain the preceding paragraph in more detail}
% \comment{tree?}
% CB: I think forest is fine here.
and
accessing neighbors only relative to the current patch leads to a high
percentage of cache reuse
\cite{BursteddeBurtscherGhattasEtAl09}.
% since the ghost patches are likely closeby with respect to the $z$-order.

@@

The threaded parallelism over the degrees of freedom of a patch can be handled
by \forestclaw alone without the need to involve \pforest.  Additionally many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for the integration of the hyperbolic system on a leaf-patch thereby allowing for hybrid parallelism.  The design of leaf-patches then can enable efficient management of data local to many-core architectures (e.g. the Xeon Phi), and the host.

>>>>

\newcommand{\comment}[1]{\textcolor{green}{[DAC: #1]}\xspace}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}\xspace}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\sR}{\mathbb{R}}
\newcommand{\dt}{\mathrm{d}t}
\newcommand{\clawpack}{{\sc Clawpack}\xspace}
\newcommand{\forestclaw}{Forestclaw\xspace}
\newcommand{\pforest}{\texttt{p4est}\xspace}
\newcommand{\manyclaw}{Manyclaw\xspace}

@@

A main challenge is presented by the fact that the patch neighborhood
is only known to \pforest.  This patch connectivity information needs
to be propagated to the numerical code in \forestclaw that implements
the interaction with neighbor patches via the use of a layer of (typically two)
ghost cells surrounding each patch.  To this end, we have designed an
interface that provides read-only access to the sequence of blocks, the list of
patches for each, and a constant-time lookup of neighbor patches (and their
relative orientation which is nontrivial between blocks).
% Since a neighbor patch has $m \ge 4$ cells in each direction, the required
% ghost cells are available by construction.
Suitably informed by \pforest, \forestclaw stores only the patches local to
each MPI rank.

Mesh modification directives, such as adaptive refinement and coarsening, are
called from \forestclaw and executed as black-box operations inside \pforest.
The only information that flows back to \pforest is a set of per-patch
refinement and coarsening flags, which are computed from the numerical state in
\forestclaw.  Our current algorithm requires neighbor patches to have a size
difference of at most a factor of two (the 2:1 balance condition).  This condition
can have nonlocal effects on the refinement pattern and is enforced by
special-purpose parallel algorithms in \pforest \cite{IsaacBursteddeGhattas12}.

@@

\forestclaw uses iterators over all rank-local leaves, optionally restricted to
a given level.  Random access is possible and used when executing $\cO(1)$ time
neighbor lookups.  Looping over the patches in the order prescribed by the
forest
\alert{Review 2 - Explain the preceding paragraph in more detail}
% \comment{tree?}
% CB: I think forest is fine here.
and
accessing neighbors only relative to the current patch leads to a high
percentage of cache reuse
\cite{BursteddeBurtscherGhattasEtAl09}.
% since the ghost patches are likely closeby with respect to the $z$-order.

\alert{The exchange of ghost} data and parallel repartitioning is
provided transparently by \pforest.
\alert{Review 2 requests more info on how this is done.  Sounds like they are more curious about the refinement overhead to other leaves on the tree and the strategy as to how refinement is decided.
%
Quote: How is AMR working in detail. For example: Figure 1 shows a
2:1 refinement, thus refinement of a single patch could affect a whole region
of the computational domain. + The underlying equations are known to
Forestclaw: How is is the information shared with p4est (equation dependent
refinement criteria, refinement and coarsening strategies (interpolation,
integration).
}

@@

The threaded parallelism over the degrees of freedom of a patch can be handled
by \forestclaw alone without the need to involve \pforest.  Additionally,
many-core implementations, such as \manyclaw \cite{manyclaw}, can be used for
the integration of the hyperbolic system on a leaf-patch thereby allowing for
hybrid parallelism.  The design of leaf-patches then can enable efficient
management of data local to many-core architectures (for example the Xeon Phi
coprocessor) and the host.
